---
title: "HW 1"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE,  
                      warning = FALSE, message = FALSE, 
                      fig.align = "center",
                      R.options = list(max.print=100))

# Edit the file starting below
```

### Enter your name and EID here:

**You will submit this homework assignment as a pdf file on Gradescope.**

*For all questions, include the R commands/functions that you used to find your answer (show R chunk). Answers without supporting code will not receive credit. Write full sentences to describe your findings.*

------------------------------------------------------------------------

## Part 1: (11 pts)

The dataset `mtcars` was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and other aspects of automobile design and performance for different cars (1973-74 models). Look up the documentation for this data frame with a description of the variables by typing `?mtcars` **in the console** pane.

### Question 1: (2 pt)

Take a look at the first 6 rows of the dataset by using an `R` function in the code chunk below. Do you know about any (or all) of these cars?

```{r}
# your code goes below (make sure to edit comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 2: (2 pts)

How many rows and columns are there in this data frame in total?

```{r}
# your code goes below (make sure to edit comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 3: (1 pt)

Save `mtcars` in your environment and name it as your `eid`. From now on, use this new object instead of the built-in dataset.

```{r}
# your code goes below (make sure to edit comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 4: (2 pts)

When is your birthday? Using indexing, grab the value of `mpg` that corresponds to the day of your birthday (should be a number between 1 and 31).

```{r}
# your code goes below (make sure to edit comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 5: (2 pts)

Using logical indexing, count the number of rows in the dataset where the variable `mpg` takes on values greater than 30.

```{r}
# your code goes below (make sure to edit comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 6: (2 pts)

Let's create a new variable called `kpl` which converts the fuel efficiency `mpg` in kilometers per liter. Knowing that 1 mpg corresponds to 0.425 kpl, complete the following code and calculate the max kpl:

```{r}
# Add a new variable to the dataset
lg28373$kpl <- lg28373$mpg 
```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

## Part 2: (6 pts)

Let's quickly explore another built-in dataset: `airquality` which contains information about daily air quality measruements in New York, May to September 1973. 

### Question 7: (2 pts)

Calculate the mean `Ozone` (in ppb). Why does it make sense to get this answer? *Hint: take a look at the column `Ozone` in the dataset.*

```{r}
# your code goes below (make sure to edit comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 8: (2 pts)

Look at the documentation for the function `mean()` by running `?mean` **in the console**. What argument should be used to find the mean value that we were not able to get in the previous question? What type of values does that argument take?

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

### Question 9: (2 pts)

Sometimes the R documentation does not feel complete. We wish we had more information or more examples. Find a post online (include the link) that can help you use that argument in the `mean()` function. Then finally find the mean ozone!

```{r}
# your code goes below (make sure to edit comment)

```

**Your answer goes here. Write sentences in bold.**

------------------------------------------------------------------------

## Formatting: (3 pts)

Knit your file! Into pdf directly or into html. 

Is it working? If not, try to decipher the error message (look up the error message, consult websites such as [stackoverflow](https://stackoverflow.com/) or [crossvalidated](https://stats.stackexchange.com/). 

Once it knits in html, click on `Open in Browser` at the top left of the window pops out. Print your html file into pdf from your browser. Any issue? Ask your classmates or TA!

------------------------------------------------------------------------

```{r, echo=F}
## DO NOT DELETE THIS BLOCK!
Sys.info()
```
